### 2015-05-11 draft.tex
minor changes

\title{Survivable Set Connectivity}
\author{
 by Fermi Ma and Erik Waingarten
}
\date{\today}

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{clrscode4e}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={210mm,297mm},
 left=20mm,
 right=20mm,
 top=20mm,
 bottom=20mm,
 }

\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}

\begin{document}
\maketitle

\section{Introduction}

In this paper, we analyze the problem of Survivable Set Connectivity (SSC), a generalization of the classical Set Connectivity problem. Set connectivity is formulated as:
\begin{quote}
\textbf{Input}: A graph weighted $G = (V, E)$, $c: E \rightarrow \mathbb{R}$ and a collection of sets $(S_i, T_i)$. \\
\textbf{Output}: A minimum cost subgraph $H \subset G$ such that there is a path between $S_i$ and $T_i$ for each $i$. 
\end{quote}
In SSC, the problem is modified so that each pair of sets $(S_i,T_i)$ must be connected by $k_i$ paths, instead of just 1. More formally, the problem is:

\begin{quote}
\textbf{Input}: A graph weighted $G = (V, E)$, $c: E \rightarrow \mathbb{R}^{\geq 0}$ and a collection of sets $(S_i, T_i)$, along with a value $k$. \\
\textbf{Output}: A minimum cost subgraph $H \subset G$ such that there are $k$ edge-disjoint paths between $S_i$ and $T_i$ for each $i$. 
\end{quote}

Unfortunately, solving exact Survivable Set Connectivity is $\mathsf{NP}$-hard \cite{ssc}. Therefore, the focus of this paper is on an approximate solution to a relaxation of this problem given in \cite{ssc}. Instead of requiring at least $k$ edge-disjoint paths between each pair, we require that each pair have at least $\Omega(\frac{k}{\log n})$ edge-disjoint paths. With this relaxation, it is possible to find a subgraph with cost $O(polylog(n))$ times the optimum.

The paper is organized as follows. We first present a natural integer linear programming formulation of the problem. Then, we introduce the idea of a \emph{tree embedding}, which allows us to transform arbitrary graphs $G$ into trees that can be more easily handled by the algorithm. We also show a general rounding scheme, called RoundGKR, which our algorithm uses as a black box. Finally, we combine all these concepts together to present and analyze an approximation algorithm for SSC. 
%
%[ GIVE MORE EXAMPLES HERE ]
%
%As with many of these kinds of problems, SSC is $\mathsf{NP}$-hard. The common route taken in 6.856 has been to find polynomial time algorithms for approximations; however, in the case with SSC, it is $\mathsf{NP}$-hard to achieve a polylogarithmic approximation. 
%
%Another possible approach is therefore to relax the question. Instead of requiring at least $k$ edge-disjoint paths, we will allow some sets to have less than $k$ paths, but not too much less. We will present an approximation algorithm which outputs a subgraph of cost $O(polylog(n))$ times the optimum such that for each $(S_i, T_i)$ pairs, there are at least $\Omega(\frac{k}{\log n})$ edge-disjoint paths between $S_i$ and $T_i$. 
%
%The approximation algorithm will follow the usual approach to approximation algorithms of this kind. We will first write the problem as an integer linear program, we solve the linear programming relaxation, and then we round the solution. 
%
%The analysis will be slightly different from the analysis shown in class. Usually, our randomized algorithm will return a feasible solution to the problem, and our analysis work is mostly to bound the cost. This was the case with max-sat, set-cover, and multi-commodity flows. In our case, we will show that the rounded solution will be close to the optimal, but we will work to show that the rounded solution is feasible.

\section{An Integer Linear Program}

In this section, we give an integer linear program that exactly solves SSC. Recall that the goal is to output a subgraph $H$ of $G$ such that each $(S_i,T_i)$ pair has at least $k$ edge disjoint paths. Naturally, we create an indicator variable $x_e$ for each edge based on whether or not it is included in $H$. 
\begin{align}
 x_e &= \left\{ \begin{array}{cc} 1 & e \in H \\
                                  0 & \text{ otherwise. } \end{array} \right.
\end{align}
Thus, our output $H$ will be the subgraph consisting of edges $e$ where $x_e = 1$. \\ \\

We claim that the following ILP captures SSC exactly. 
SSC-LP:
\begin{align}
\min & \sum c(x) x_e  \\
\text{ s.t } & \sum_{e \in \delta(S)} x_e \geq k, \text{for each }S\text{ where } S_i \subset S,T_i \subset V - S, \\ & \hspace{1in}\delta(S) \text{ is the edges of the $S$ cut } \\
& x_e \in [0,1] \text{ for all }e \in E
\end{align}

The objective function $\sum c(x) x_e$ simply minimizes the cost of all edges included in $H$. The constraint ensures that for all pairs $(S_i,T_i)$, any cut where $S_i$ is on one side of the cut and $T_i$ is on the other side has at least $k$ edges crossing the cut. This property holds if and only if there are $k$ edge disjoint paths from $S_i$ to $T_i$. 
%
%
% 
%This cut-based integer linear program finds exactly $H$. To see why the constraints ensure that each $(S_i, T_i)$ pair has at least $k$ edge disjoint paths, we can look at the cuts corresponding to $S_i$ and $T_i$. First, we let $S = S_i$, then there are $k$ outgoing edges. We can choose to include one more node to $S$, following a particular outgoing edge of the cut in $H$. We can continue doing this until we reach $T_i$. This will trace out a path between $S_i$ and $T_i$. If we do this for each edge, and then we do this for each $(S_i, T_i)$ pair, we can confirm that there are in fact $k$ edge-disjoint paths between each $(S_i, T_i)$ pair.

Of course, since this is an integer linear program, we must relax it to fractional solutions in order to solve it. Thus, we relax the constraint of $x_e \in [0,1]$ to $0 \leq x_e \leq 1$. However, there is still a slight issue; there can be exponentially many constraints. This is because there are exponentially many choices of $S$ for each $(S_i,T_i)$ pair. To get around this issue, we note that the ellipsoid algorithm can handle linear programs with arbitrarily many constraints, as long as it is given a separation oracle. Luckily, we can give such an oracle.

The separation oracle just computes the minimum $S_i, T_i$ cut in the graph. What we can do is contract $S_i$ and $T_i$ to single vertices, and then compute the min-cut in the remaining graph. If there is a min-cut of size less than $k$, then that cut will correspond to a violating contraint. If the min-cut has size at least $k$, then all cuts have size at least $k$, so we have satisfied the constraints.

%The next tool we will need are tree embeddings. We have seen metric embeddings in class. Basically, we had a desired type of object $O$ (in the case of cuts, we had a metric space corresponding to a cut) what we did was come up with a metric space $(X, d)$ and a map $f: (X, d) \rightarrow O$ such that the map satisfied certain properties on the distances. 
%
%Here we will have a similar goal. We want to understand minimum cost weighted graphs, but since analyzing them is hard, we instead study minimum cost weighted trees, and then we map the trees back to the graphs. 
%
%\begin{definition}
%A tree embedding of a weighted graph $G = (V, E)$ will be a tree $T = (V_T, E_T)$ with weights $y: E \rightarrow \mathbb{R}^{\geq 0}$ and a function $\text{map} : V_T \cup E_t \rightarrow V \cup P(E)$ ($P(E)$ is the power set of $E$) such that
%\begin{align}
%\text{map}(V_t) &\subset V \\
%\text{map}(E_t) &\subset E
%\end{align}
%In other words, vertices map to vertices and edges map to edges. We will usually refer to an embedding as $(T, y, \text{map})$. 
%\end{definition}

\section{R\"{a}cke's Trees}

We will use a particular type of tree embedding given in \cite{racke} which will allow us to map paths in the tree to paths in the graph.

\begin{definition}
For a given graph $G = (V, E)$ with edge capacities $x_e \geq 0$, $e \in E$, a R\"{a}cke's Tree embedding is a triple of the form $(T,y,\mathrm{map})$ where $T = (V_T,E_T)$ is a rooted tree, $y$ is a capacity function on the edges $E_T$ of $T$, and $\mathrm{map}$ is a function $V_T \cup E_T \to V \cup 2^E$ satisfying the following properties:
\begin{enumerate}
\item For each node $v \in V_T$, $\mathrm{map}(v) \in V$. Thus, the $\mathrm{map}$ function maps nodes in the tree $T$ to nodes in the original graph, $G$.
\item Furthermore, $\mathrm{map}$ restricted to the leaves of $T$ is a bijection with the vertices of $G$. In other words, each leaf of the tree $T$ is in one-to-one correspondence with a vertex of the original graph, $G$.
\item For each edge $e = (u,v) \in E_T$, $\mathrm{map}(e)$ is a set of edges that form a path between $\mathrm{map}(u)$ and $\mathrm{map}(v)$. Note that $\mathrm{map}(u)$ and $\mathrm{map}(v)$ are vertices in $G$, so $\mathrm{map}(e)$ specifies a set of edges in $G$.
%\item For some edge $e \in E_T$, $y_e$ (the cost of the edge) is a special quantity. Removing $y_e$ cuts the tree into two components, and so the leaves in the two components form a cut $\delta(S)$. Figure~\ref{fig:rackecut} gives an example for one such cut induced by an edge.
%\begin{align}
%y_e &= \sum_{e' \in \delta(S)} x_e
%\end{align}
\end{enumerate}
\end{definition}

With this setup, we can define the inverse function, $\mathrm{map}^{-1}$. If we try to compute $\mathrm{map}^{-1}(v)$ for some $v \in V$, we note that multiple vertices in $T$ could map to $v$. Thus, we specify that $\mathrm{map}^{-1}(v)$ be the unique \emph{leaf} vertex in $V_T$ that maps to $v$. We will also consider $\mathrm{map}^{-1}$ applied to sets of vertices $W \subseteq V$, and we define it in the obvious way: $\mathrm{map}^{-1}(W) = \bigcup_{v \in W} \mathrm{map}^{-1}(v)$.

For edges $e = (u,v) \in E$, we define $\mathrm{map}^{-1}(e) = \{ f \in E_T | e \in \mathrm{map}(f)\}$. This is simply the set of all edges $f$ in the tree $T$ that get mapped to a set of edges $\mathrm{map}(f)$ that contain $e$. Again, this defintiion can be extended to sets of edges $F \subseteq E$, so that $\mathrm{map}^{-1}(F) = \bigcup_{e \in F} \mathrm{map}^{-1}(e)$.

Finally, for an edge $e \in E_T$, we define $y_e$ as follows. The removal of $e$ from $T$ separates the tree into two connected components. Let $S$ be the set of vertices of $G$ mapped to by the leaves of one of the connected components, and so $V \setminus S$ is the set of vertices of $G$ mapped to by the leaves of the other connected component. Define $\delta(S)$ to be the set of edges in $E$ that cross from $S$ to $V \setminus S$. Then we set $y_e = \sum_{e' \in \delta(S)} x_{e'}$, the total edge capacity crossing the cut.

\begin{figure}
\centering
\label{fig:rackecut}
\includegraphics[width=0.5\linewidth]{Rackecut.pdf}
\caption{R\"{a}cke's tree cut induced by an edge in the tree.}
\end{figure}

\begin{definition}
Given a R\"{a}cke tree, the load on an edge of the graph $e \in E$ is
\begin{align}
load(e) &= \sum_{f \in \text{map}^{-1}(e)} y_f,
\end{align}
and the relative load is defined as 
\begin{align}
rload(e) = \dfrac{load(e)}{x_e}.
\end{align}
\end{definition}

The usefulness of R\"{a}cke's trees becomes apparent when we consider multi-commodity flows. Recall that the multi-commodity flow problem specifies $k$ different commodities, where each commodity $i$ must be routed from source $s_i$ to sink $t_i$, subject to constraints on the capacity of each edge. The resulting flow can be represented as a set of functions $\{ f_i \}$, where $f_i(e)$ is the nonnegative real number amount of commodity $i$ flowing on edge $e$.

There is a natural way to map multi-commodity flows between $G$ and $T$. In particular, suppose we have a multi-commodity flow $\{ f_i \}$ on $G$. For each $i$, if there are $f_i$ units of flow going from $s_i$ to $t_i$ in $G$, we route $f_i$ units of flow in the $T$ from $\mathrm{map}^{-1}(s_i)$ to $\mathrm{map}^{-1}(t_i)$. If we are given a multicommodity flow $\{ f_i^T \}$ in $T$, then if we have $f_i^T(e)$ units of flow on edge $e$ in $T$, then we route $f_i^T(e)$ units of flow through the path $\mathrm{map}(e)$ in $G$ (recall that edges in $T$ map to paths in $G$).

Figure~\ref{fig:racketree} gives an example of the flow correspondence between a graph and it's associated R\"{a}cke's tree. Each red vertex of the tree is mapped to another red vertex on the graph, and the edge between them corresponds to a path between the red vertices. This gives us a simple way to map flows from Racke trees to the graph and vice-versa.

\begin{figure}
\label{fig:racketree}
\centering
\includegraphics[width=0.8\linewidth]{Rackeembeddinggraph.pdf}
\caption{Example Racke Tree where red nodes are mapped and edge maps to a path.}
\end{figure}

There are some crucial properties of Racke trees which we will treat as black boxes. We will define these as lemmas without proof, and we refer to [CITE] for the proofs.

\begin{lemma}[\cite{racke}]
\label{lem:mapflows}
If we have a graph $G = (V, E)$ and capacities $x: E \rightarrow \mathbb{R}^{\geq 0}$ and a Racke tree $(T, y, \text{map})$. A multicommodity flow on $G$, $\{ f_i \}$, maps directly to a multicommodity flow on the tree which is also feasible. Likewise, a multicommodity flow on $T$ maps directly into a multicommodity flow on the graph, with the small caveat that we must multiply the capacity of the edge by the relative load. 
\end{lemma}

\begin{lemma}[\cite{racke}]
\label{lem:rload}
For some given graph $G = (V, E)$ and capacities $x: E \rightarrow \mathbb{R}^{\geq 0}$, we can sample efficiently from Racke trees of $G$, such that 
\begin{align}
\max_{e \in E} \textbf{E}[rload(e)] &\leq \alpha = O(\log n) 
\end{align}
\end{lemma}

\begin{lemma}[\cite{ssc}]
\label{thm:height}
The height of the trees sampled is $O(\log nC)$, where $C$ is the ratio of the largest to smallest capacity in $G$. In our cases, the height will always be $O(\log n)$. 
\end{lemma}

\section{The RoundGKR Procedure}

The RoundGKR algorithm is a rounding procedure used to solve the Group Steiner Tree (GST) problem \cite{GKR}. The problem it stated as follows. \\

Group Steiner Tree:
\begin{quote}
\textbf{Input}: A tree $T = (V_T, E_T)$ rooted at $r$ where each edge has a nonnegative real value cost $c(e)$, along with a set $\{T_i\}$ where each $T_i \subseteq V_T$.\\
\textbf{Output}: The minimum cost subtree $H \subseteq T$ with a path from $r$ to $t_i$ for some $t \in T_i$ for all $i$.
\end{quote}

The following integer linear program solves the problem exactly:\\
GST-ILP:
\begin{align}
\min & \sum c(e) y_e  \\
\text{ s.t } & \sum_{e \in \delta(S)} y_e \geq 1, \text{ where } S_i \subset S, T_i \subset V - S, \delta(S) \text{ is the edges of the $S$ cut }
\end{align}

If we relax the integer linear program to a linear program and solve it, we can use RoundGKR to round the resulting fractional solution to an integer solution. The optimum to the linear program will assign fractional values in $[0,1]$ to each $y_e$. For any $e$, let $p(e)$ denote its parent edge (if it exists). We claim that since our problem is to find a subtree that includes the root $r$, edges that are closer to the root are at least as likely to be included in the subtree than edges farther from the root. Since the value of $y_e$ will be used as the probability that edge $e$ is included in $H$, we will assume that $y_{p(e)} \geq y_e$.

The RoundGKR procedure simply starts at the root $r$ and builds the subtree $H$ by working its way down the original tree, $T$. It decides whether or not to include edge $e$ only once it has included edge $p(e)$ (if ever). At this step, it includes $e$ with probability $\frac{y_e}{y_{p(e)}}$. A simple inductive argment shows that an edge $e$ is chosen with probability $y_e$ (proof sketch: assume by induction that we know $Pr[\text{include $p(e)$}] = y_{p(e)}$. Then $Pr[\text{include $e$}] = Pr[\text{include $e$ $|$ include $p(e)$}]*Pr[\text{include $p(e)$}] = \frac{y_e}{y_{p(e)}}y_{p(e)} = y_e$.)

The important property of the algorithm can be summarized in the following lemma, whose proof can be found in [CITE].

\begin{lemma}[\cite{GKR}]
\label{lem:findpath}
Suppose that $T = (V_t, E_t)$ support a flow of at least $1$ from $T_i$ to the root $r$ without the edges $F_t$. Then the RoundGKR returns a solution with a path in $T$ from $T_i$ to $r$ without using $F_t$ with probability at least $\Omega(\frac{1}{\log n})$. 
\end{lemma}

We will rely on this lemma heavily. In fact, since we need $\Omega(\frac{k}{\log n})$ edge-disjoint paths, we will argue that if we let $F_t$ be any set of at most $\Omega(\frac{k}{\log n})$ edges, then we still have a path, and therefore, we have $\Omega(\frac{k}{\log n})$ paths.

\section{Algorithm}

Now we will combine the ideas of the linear program to approximate a solution to Survivable Set Connectivity. We copy the algorithm as it is written in the paper below, although we will separate the algorithm into parts and explain each part in detail.
\begin{codebox}
\li Set $H \leftarrow \emptyset$
\li Compute linear program SSC-LP and set to zero any $x_e < \dfrac{k}{n^2}$.
\li Let $x' \leftarrow x*\frac{4}{k}$. 
\li Prepare $\mathcal{D}$ be the distribution of Racke tree embeddings for $G$ with $x'$.
\li \For $\tau$ rounds \Do
\li Sample $(T, y, \text{map})$ from $\mathcal{D}$. 
\li \For $q = 0, ..., 2\log n$ \Do
\li \For $v \in V_t$ \Do
\li With probability $\min\{1, \dfrac{4}{2^q}\}$, $q$-mark $v$
\li \If $v$ is $q$-marked \Then
\li Set $y^{v, q}\leftarrow 2^{q+1}y$ and $E^{v,q} \leftarrow \emptyset$.
\li \For $\tau'$ iterations, \Do
\li $E^{v,q} \leftarrow E^{v,q} \cup$ RoundGKR$(T_v, v, y^{v,q})$ \End
\li $H \leftarrow H \cup \text{map}(E^{v,q})$ \End \End \End \End
\li return $H$
\end{codebox}

\subsection{Algorithm Idea}

Here, we provide an intuitive explanation of the algorithm as well as the main motivations for each procedure. Of course, we will revist these ideas in the analysis.

After running the linear program, we rescale $x$ to $x'$ by a $\frac{4}{k}$ factor, which ensures that we have a flow of at least $4$ in each $(S_i, T_i)$. This flow of $4$ will correspond to a flow of $4$ in the R\"{a}cke's trees.

The RoundGKR algorithm is the main tool used for our rounding scheme. RoundGKR finds paths in trees which avoid specific edges with probability $\Omega(\frac{1}{\log n})$ (this is Lemma~\ref{lem:findpath}) given there is sufficient flow. We repeat this $\tau' = O(\log n)$ times to assure a constant probability of finding such paths.

However, we can only run RoundGKR on trees, not general graphs. We introduce the R\"{a}cke's tree embedding specifically for this purpose. We know a flow of $4$ exists between leaves (Lemma~\ref{lem:mapflows}). Each flow path will go up the tree to some ancestor $v$ and then back down to a leaf. One such path might be found when $v$ gets marked. One path will go to $s_i'$ and one to $t_i'$.

Lemma~\ref{lem:findpath} requires us to have a flow of at least $1$ avoiding some edges. If a flow path from $s_i'$ to $t_i'$ goes through ancestor $v$ and carries a flow of value $f$, we can apply the lemma if $f \geq 1$. We scale the capacities by $2^{q+1}$ in order to search for the flow value. We only scale with probabilities $\dfrac{4}{2^q}$ so our costs does not increase by more than $O(\log n)$. Figure~\ref{fig:roundgkronnode} shows how a flow path might be found.

\section{Algorithm Analysis}

\subsection{Overview}

We claim that the algorithm finds a solution which has
\begin{itemize}
\item (Claim A) at least $\Omega(\frac{k}{\log n})$ edge disjoint paths between each $(S_i, T_i)$ pair
\item (Claim B) cost within a polylogarithmic factor of the optimum
\end{itemize}

We first prove Claim A, and then using Claim A, we prove Claim B.

\subsection{Claim A}

\begin{definition}
For some set of edges $F \subseteq E$, the sampled tree embedding, $(T, y, \text{map})$ is bad if $F_T = \text{map}^{-1}(F)$ 
\[ \sum_{e' \in F_T} y_{e'} \geq \frac{1}{2} \]
And good otherwise.
\end{definition}

Intuitively, bad sets correspond to sets which are necessary in the graph. We want graphs to be good with respect to sets of size $\Omega(\frac{k}{\log n})$ in order to ensure connectivity.

\begin{lemma}[\cite{ssc}]
\label{lem:notbad}
Fix some $F$ of size at most $\dfrac{k}{16 \alpha}$. Then the probability that a sampled tree embedding $(T, y, \text{map})$ from $\mathcal{D}$ is bad is at most $\frac{1}{2}$. 
\end{lemma}

\begin{proof}
Suppose we have a bad embedding for the set $F$. Then 
\begin{align}
\sum_{e \in F} load(e)&= \sum_{e \in F} \sum_{e' \in \text{map}^{-1}(e)} y_{e'} \\
                    &\geq \sum_{e' \in F_T} y_{e'} \\
                    &\geq \frac{1}{2}
\end{align}
Where the first inequality we get from taking the sum over the union versus the disjoint union, and the second inequality from the assumption that the tree embedding is bad.

Therefore, a bad tree means that the load of the set of $F$ is more than $\frac{1}{2}$. Therefore,
\begin{align}
Pr[\text{pick a bad tree for $F$}] &\leq Pr[ \sum_{e\in F} load(e) \geq \frac{1}{2}]
\end{align}
From Lemma~\ref{lem:rload}, we know that
\begin{align}
\textbf{E}[ \sum_{e \in F} load(e) ] &\leq \sum_{e \in F} \textbf{E}[rload(e)]x_{e'} \\
                                   & \leq \sum_{e\in F} \alpha \frac{4}{k} x_{e} \\
                                   & \leq \dfrac{4\alpha}{k} |F| \\
                                   &\leq \frac{1}{4}
\end{align}
Since we limited the size of $F$. Now invoking Markov's inequality, we know that the probability of bad event is at most $\frac{1}{2}$. 
\end{proof}

\begin{lemma}[\cite{ssc}]
With probability at least $\frac{3}{4}$, all set pairs are at least $\frac{k}{16\alpha} + 1$ edge connected in $H$. 
\end{lemma}

\textbf{Proof Idea}:
\begin{enumerate}
\item Set a particular set of edges $F$ of size at most $\Omega(\dfrac{k}{\log n})$. 
\item Argue that with probability at most $\frac{1}{2}$, the edges in $F$ are not necessary for connectivity by claiming that the total $load$ on the edges in low. (This is Lemma~\ref{lem:notbad}). 
\item We will condition on this probability to show that if we remove $\text{map}(F)$ on the tree, we still have sufficient flow from each $s_i$ to $t_i$. This flow must go up the tree to some common ancestor $v$ and then back down.
\item Then we will claim that at some $v$, it gets marked and the RoundGKR algorithm which is rooted at $v$ finds the two paths from $v$ to $s_i$ and $t_i$, and so the path will be included with sufficient probability and exclude $F$.
\item We finally union bound over all $F$ and over all $i$ to show that we succeed with constant probability. 
\end{enumerate}

\begin{proof}
We fix an $F$ of size at most $\frac{k}{16\alpha}$ and an $i$ and show that $(S_i, T_i)$ are connected with an edge outside of $F$. I will show this happens with a probability at least $\beta$. So the probability of failure is $(1-\beta)$. If we repeat this $\tau = \frac{1}{\beta}\ln(4hn^{\frac{k}{16\alpha}})$ times, then our failure probability goes down to $\dfrac{1}{4hn^{\frac{k}{16\alpha}}}$. So now we take the union bound over all of $F$ and over all $i$, so we get probability of failure at most $\frac{1}{4}$.

It remains to show that the probability that we find a path from $S_i$ to $T_i$ without using the edges from $F$ is at least $\beta$. We do this in the following lemma.
\end{proof}

\begin{lemma}
Fix some $F$ and some $i$, where $|F| < \dfrac{16k}{\alpha}$. The probability that at round $t$ we find a path from $S_i$ to $T_i$ without using edges from $F$ is at least some constant $\beta > 0$. 
\end{lemma}

\begin{proof}
We condition on the probability that $F$ is good at round $t$, which by Lemma~\ref{lem:notbad} is at least $\frac{1}{2}$. Now we can set the $y$ terms of $F_T = \text{map}^{-1}(F)$ to zero, in order to avoid picking an edge in $F$.

Since we have a good embedding with respect to $F$. The assumption was that $G$ can support a flow of at least $k$ in the original graph (by the solution to the fractional linear program). Recall that we reweighed the values $x$ by $\frac{4}{k}$, so we can support a flow of at least $4$ in the original graph with $x'$. So by Lemma~\ref{lem:mapflows}, $T$ supports a flow from $S_i' = \text{map}^{-1}(S_i)$ to $T_i' = \text{map}^{-1}(T_i)$ of value at least $4$.

Now for some reason, when we remove a total capacity of $\frac{1}{2}$, we are left with a flow of $2 - \frac{1}{2} = \frac{3}{2}$. In particular, we can support a unit flow from $S_i'$ to $T_i'$. That unit flow can be decomposed into paths $P_1, ..., P_z$, disregarding paths with very small amount of flow $(\leq \frac{1}{2n^2})$. So each path goes up from a leaf, turns at some $v$, and then goes down to another leaf. Let $\phi_v$ be the flow turning at $v$.

Since we support a flow of at least $\frac{3}{2}$ and we removed paths with at most $\frac{1}{2n^2}$ flow, we are left with
\[ \phi_v \geq \frac{3}{2} - \dfrac{1}{2n^2}|E_T] \geq 1 \]
Now $\phi_v$ falls between some $(2^{-q_v-1}, 2^{-q_v}]$ for some $q_v \in \{0, ..., 2\log_2 n\}$. Now the probability that none of the $v \in V_T$ is $q$-marked exactly at $q_v$ when we have a good embedding is at most $e^{-3/4}$ which follows by a Chernoff bound. So conditioning on this happening, the path gets multiplied by $2^{q_v+1}$ and so it supports a flow of $1$. Now Lemma~\ref{lem:findpath} says the path will be included with probability at least $1 - (1 - \Omega(1/\log n))^{\tau'}$. If we let $\tau' = O(\log n)$, then we get any constant probability $\epsilon$. Since we need to find both paths from the root to both leaves, we get a failure probability of $2\epsilon$.

So the probability that we find a path outside of $F$ is at least one minus the union bound over failure probabilities. This is $1 - \frac{1}{2} - e^{-3/4} - 2\epsilon > 0$. 
\end{proof}

\subsection{Claim B}

\begin{lemma}[\cite{ssc}]
$\textbf{E}[c(H)] = O(\log^4n(\frac{\log h}{k}+1)) * c(x)$, where $c(x) = \sum_e c(e)x_e$. 
\end{lemma}

\begin{proof}
We prove this by bounding the probability that a given edge is included in $H$ by $O(\log^4n(\frac{\log h}{k} + 1))x_e$. The claim then follows by linearity of expectation.

Let $(T = (V_t, E_t), y, \text{map})$ be a Racke tree sampled from $\mathcal{D}$. For an edge $e$ to be included in our output subgraph $H$, there must have been some tree edge $e' \in E_T$ that maps to $e$ (so $e \in \mathrm{map}(e')$) selected to be in $E^{v,q}$ for some vertex $v \in V_t$ and some index $q \in [0,\dots,2\log n]$. For this to occur, $v$ must be $q$-marked, an event that occurs with probability $\frac{4}{2^{-q}}$.

For a given edge $e'$, there are at most $height(T)$ ancestors, which means that there are at most $height(T)$ different vertices $v$ that, if $q$-marked, may lead to $e'$ being included. For each $v$, we must consider all $O(\log n)$ possible values of $q$. For each value of $q$, the probability of being $q$-marked is $\frac{4}{2^{-q}}$. Once the vertex is $q$-marked, we have $\tau'$ trials of the RoundGKR algorithm, where in each trial, there is a $2^{q+1}y_{e'}$ chance of edge $e'$ being included (by construction of the algorithm). Multiplying these probablities together and applying a union bound over all $\tau'$ trials and over all possible values of $q$, we get an upper bound of

\begin{align}
height(T) \sum_q \frac{4}{2^{-q}} 2^{q+1} y_{e'} \tau' &= height(T) \tau' y_{e'} \sum_q 2 = O(\log^3 n) y_{e'} 
\end{align}

So the probability that a given edge of the tree is included in a particular sampling of a Racke tree is $O(\log^3 n) y_{e'}$.

So now in order to include some edge $e$ of the graph $G$, we must include some edge $e' \in \text{map}^{-1}(e)$ (recall that this is the set of tree edges which, if selected, result in edge $e$ being included). So let $X_e$ be the indicator random variable which says that edge $e$ is included in $H$, and $Y_{e'}$ be the indicator random variable which says that edge $e'$ in the tree is included in some RoundGKR execution. Above, we argued that $E[Y_{e'}] \leq O(\log^3 n)y_{e'}$.
\begin{align}
X_e \leq \sum_{e' \in \text{map}(e)} Y_{e'}
\end{align}
Because if $e$ is included in the graph, it must have come from some edge in the tree. 
\begin{align}
 \textbf{E}[X_e] &\leq \tau \sum_{e' \in \text{map}^{-1}(e)} O(\log^3 n) y_{e'} \\
        &= O(\log^3 n) \tau \textbf{E}[load(e)] \\
        &= O(\log^3 n) \tau \textbf{E}[rload(e)x_e'] \\
        &\leq O(\log^3 n) \tau \dfrac{4}{k} \alpha x_e\\
        & = \dfrac{O(\log^4 n)\tau}{k} x_e
\end{align}
So now, bounding the optimum, we get
\begin{align}
\textbf{E}[c(H)] &= \sum_{e \in E} c(x_e)\textbf{E}[X_e] \\
                 &\leq \dfrac{O(\log^4 n)\tau}{k} \sum_{e \in E} c(x_e) x_e
\end{align}
So when $\tau = O(\log h + k)$, we get our desired approximation.
\end{proof}

\begin{figure}
\label{fig:roundgkronnode}
\centering
\includegraphics[width=0.6\linewidth]{RoundGKRonNode.pdf}
\caption{Execution of GKR on some subtree rooted at $v$ in a particular round finding paths to $s_i$ and $t_i$.}
\end{figure}

Having established that the expected cost is only a polylogarithmic factor away from the optimum, it remains to show that the edges chosen do in fact form a graph which connects $(S_i, T_i)$ with $\Omega(\frac{k}{\log n})$ disjoint paths for each $i$. We will need the following definition.

\bibliography{references}
\bibliographystyle{splncs}

\end{document}